Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-08-26 idp/documentation/paper.tex
fixed some formatting, added definitions

@@

			Note that there may be multiple ways to describe a given curve design tool in terms of this process. Instead of modelling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify. For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve. Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + (p_1 - p_0) \cdot t\), thus uniquely identifying the desired curve. Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool. In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(G^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).

@@

				Bézier splines are probably the most widely used type of curve in two-dimensional vector graphics. Bézier splines are defined piecewise using Bézier curves, which are planar parametric curves (functions \(\left[0,1\right] \rightarrow \mathbb{R}^2\)) realized using polynomials (usually cubic) in Bernstein form. A Bézier curve \(B\) of degree \(n\) consists of \(n + 1\) coefficients \(P_i \in \mathbb{R}^2\). The curve will always go through the first and the last coefficient point (\(B\left(0\right) = P_0, B\left(1\right) = P_n\)) and stay within the convex hull of the coefficient points at all other times. Bézier curves in a Bézier spline are usually pieced together such that the resulting spline is \(G^1\) continuous.

@@

				Spiro splines consist of parts of the Euler spiral, which is a curve whose curvature changes linearly with the curve's arc length. The user can specify control points through which the software puts an interpolating spline. There is no way to directly specify tangent direction and/or curvature. The splines are guaranteed to have \(G^2\) continuity.

@@

				The only real issue lies in the insufficient expressiveness of the specification language. Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability. Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature. These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature. This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature. Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties. Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified. These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved. That way, when indirectly specifying direction and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.

@@

				Unfortunately, as can be seen in example 3, this scenario does not handle modifications very well. For instance, increasing the curvature of segment c has to be done indirectly by moving a control point. Unfortunately, in this case, this also changes the direction of the curve at that point, such that additional modifications would be needed to correct this. The change also adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.

@@

@@

				The specification language we propose uses differential geometric properties of curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use. The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will use points on the curve (denoted \(p\)), tangent angles (denoted \(\theta\)) of the curve as well as the curvature value (denoted \(\kappa\)) of the curve. In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties.

				With point specifications, it's sometimes enough to require the curve to go through all of them in order, although sometimes, this will lead to undesirable results. Even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points. When specifying tangent angles and/or curvature values without an associated point, it becomes a necessity to associate this specification with some kind of positional information in order to anchor it to some point on the curve. If this was not done, the specification could move freely along the curve and would thus behave very unpredictably. We thus chose to associate each combination of point, tangent angle and curvature value with a position on the curve.

				With that in mind, it is now possible to abandon this notion of a combination of point, tangent angle and curvature value and instead talk about specification items, a specification item being either a positioned point, a positioned tangent angle or a positioned curvature value. A combination of point and curvature value at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.

@@

				Many different fairness measures for plane curves have been proposed throughout history. Most of them set out to emulate the behavior of mechanical splines made from wood or steel, like those which were used for industrial design before the advent of computer aided design. The most direct approach of this type uses mechanics to derive a term for the energy in a bent steel band, arriving at the following functional describing the minimum energy curve (MEC):
				\begin{displaymath}
					\int_{0}^{1}{\kappa\!\left(t\right)}^2\;\mathrm{d}t
				\end{displaymath}

				While this approach is a good idealization of the mechanical spline, it has some disadvantages when used for curve design. For instance, it will not yield a circular arc when constrained by co-circular points. Thus, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves. One such approach is that of the minimum variation curve (MVC), whose functional is given below:
				\begin{displaymath}
					\int_{0}^{1}{\kappa'\!\left(t\right)}^2\;\mathrm{d}t
				\end{displaymath}

				Minimum variation curves capture the intuitive notion of smoothness very well. The functional will put straight lines through points that are co-linear and circular arcs through points that are co-circular when otherwise unconstrained. When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvture, since that will make the derivation of the curvature a constant and thereby minimize the integral over the square fo the derivative of the curvature. Because of these appealing properties, the MVC fairness measure was chosen.

@@

				definitions 
				\begin{align*}
					p & : \left[0,1\right] \rightarrow \mathbb{R}^2 & & && \text{(point)}\\
					v & : \left[0,1\right] \rightarrow \mathbb{R}^2, & v & = p' && \text{(velocity)}\\
					a & : \left[0,1\right] \rightarrow \mathbb{R}^2, & a & = v' && \text{(acceleration)}
				\end{align*}

				The first problem we face is that polynomials will in general not be in natural parametrization (that is, a parametrization where the parameter \(t\) is equal to the covered arc length of the curve up to the point \(p\left(t\right)\)). Since we chose to give the position of a specification item as a fraction of the curve's total arc length, it would be very advantageous to have a parametrization where the parameter is at least proportional to the covered arc length of the curve. This property is equivalent to the first derivative of the curve (velocity) having constant euclidean norm (\(\|p'\left(t\right)\|\) is constant) and has additional useful consequences, such as the second derivative of the curve (acceleration) being orthogonal to the velocity. Thus, one of the objectives of the optimization is to make sure that the speed of the curve is constant.

>>>>

% math commands
\DeclareMathOperator{\atan2}{atan2}
\newcommand{\apply}[2]{#1\!\left(#2\right)}
\renewcommand{\vector}[2]{\begin{pmatrix}#1\\\linebreak{}#2\end{pmatrix}}

@@

			Note that there may be multiple ways to describe a given curve design tool in terms of this process. Instead of modelling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify. For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve. Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \left(p_1 - p_0\right) \cdot t\), thus uniquely identifying the desired curve. Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool. In a similar fashion, one may consider shortcomings (for instance insufficient continuity) as either shortcomings of the specification language (the language does not allow the user to request \(G^2\) continuity) or as shortcomings of the fairness measure (the fairness measure does not select sufficiently smooth curves).

@@

				Bézier splines are probably the most widely used type of curve in two-dimensional vector graphics. Bézier splines are defined piecewise using Bézier curves, which are planar parametric curves (functions \(\left[0,1\right] \rightarrow \mathbb{R}^2\)) realized using polynomials (usually cubic) in Bernstein form. A Bézier curve \(B\) of degree \(n\) consists of \(n + 1\) coefficients \(P_i \in \mathbb{R}^2\). The curve will always go through the first and the last coefficient point (\(\apply{B}{0} = P_0, \apply{B}{1} = P_n\)) and stay within the convex hull of the coefficient points at all other times. Bézier curves in a Bézier spline are usually pieced together such that the resulting spline is \(G^1\) continuous.

@@

				Spiro splines consist of parts of the Euler spiral, which is a curve whose curvature changes linearly with the curve's arc length. The user can specify control points through which the software puts an interpolating spline. There is no way to directly specify tangent angle and/or curvature. The splines are guaranteed to have \(G^2\) continuity.

@@

				The only real issue lies in the insufficient expressiveness of the specification language. Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability. Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify tangent angle nor curvature. These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of tangent angle and/or curvature. This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature. Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties. Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified. These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved. That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.

@@

				Unfortunately, as can be seen in example 3, this scenario does not handle modifications very well. For instance, increasing the curvature of segment c has to be done indirectly by moving a control point. Unfortunately, in this case, this also changes the tangent angle of the curve at that point, such that additional modifications would be needed to correct this. The change also adversely affects segment a, since the curvature there was not specified directly but instead relied on the rather fragile configuration of the 4 control points.

@@

		\subsection{Parametric Plane Curves}
		\label{section:parametric_plane_curves}

			In order to be able to reason about curves, a few basic definitions are needed.

			Definitions about vectors in the euclidean plane:
			\begin{align*}
				\overline{\cdot} & : \mathbb{R}^2 \rightarrow \mathbb{R}^2     & \overline{v}      & = \vector{-v_2}{v_1}                      && \text{orthogonal vector}\\
				\alpha           & : \mathbb{R}^2 \rightarrow \mathbb{R}       & \apply{\alpha}{v} & = \apply{\atan2}{v_2, v_1}                && \text{polar coordinate angle}
			\end{align*}

			Definitions about parametric curves:
			\begin{align*}
				p                & : \left[0,1\right] \rightarrow \mathbb{R}^2 &                   &                                           && \text{point}\\
				s                & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{s}{t}      & = \left\|\apply{p'}{t}\right\|            && \text{speed}\\
				\theta           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\theta}{t} & = \apply{\alpha}{\apply{p'}{t}}           && \text{direction}\\
				\kappa           & : \left[0,1\right] \rightarrow \mathbb{R}   & \apply{\kappa}{t} & = \frac{\apply{\theta'}{t}}{\apply{s}{t}} && \text{curvature}
			\end{align*}

			Using these definitions, the direction \(\apply{\theta}{t}\) is equal to the tangent angle at the point \(\apply{p}{t}\) and the curvature \(\apply{\kappa}{t}\) is equal to the inverse of the radius of the osculating circle at the point \(\apply{p}{t}\) or \(0\) if the curve is locally a straight line at the point \(\apply{p}{t}\).

@@

				The specification language we propose uses differential geometric properties of curves and is thus decoupled from specifics of the underlying curve model, hopefully leading to a specification language that is easier to learn and use. The human visual system seems to be fairly insensitive to curve properties beyond the second derivative, so the specification language will focus on the properties point, direction and curvature as defined in \ref{section:parametric_plane_curves}. In order to provide the maximum flexibility, the specification language should allow to specify any combination of these three properties.

				With point specifications, it's sometimes enough to require the curve to go through all of them in order, although sometimes, this will lead to undesirable results. Even with the curve preserving the order in which the points are visited, it is unspecified how much arc length the curve will spend between each pair of points. When specifying direction and/or curvature without an associated point, it becomes a necessity to associate this specification with some kind of positional information in order to anchor it to some point on the curve. If this was not done, the specification could move freely along the curve and would thus behave very unpredictably. We thus chose to associate each combination of point, direction and curvature with a position on the curve.

				With that in mind, it is now possible to abandon this notion of a combination of point, direction and curvature and instead talk about specification items, a specification item being either a positioned point, a positioned direction or a positioned curvature. A combination of point and curvature at some position thus splits up into a point specification item at this position and a curvature specification item at the same position.

@@

				Many different fairness measures for plane curves have been proposed throughout history. Most of them set out to emulate the behavior of mechanical splines made from wood or steel, like those which were used for industrial design before the advent of computer aided design. The most direct approach of this type uses mechanics to derive a term for the energy in a bent steel band, arriving at the following functional describing the minimum energy curve (MEC):
				\begin{displaymath}
					\int_{0}^{1}\apply{\kappa}{t}^2\;\mathrm{d}t
				\end{displaymath}

				While this approach is a good idealization of the mechanical spline, it has some disadvantages when used for curve design. For instance, it will not yield a circular arc when constrained by co-circular points. Thus, there has also been some research into fairness measures whose design is more directly motivated by the human perception of smoothness in plane curves. One such approach is that of the minimum variation curve (MVC), whose functional is given below:
				\begin{displaymath}
					\int_{0}^{1}\apply{\kappa'}{t}^2\;\mathrm{d}t
				\end{displaymath}

				Minimum variation curves capture the intuitive notion of smoothness very well. The functional will put straight lines through points that are co-linear and circular arcs through points that are co-circular when otherwise unconstrained. When constraints cause different points on the curve to have different curvatures, the functional will prefer linearly changing curvture, since that will make the derivation of the curvature a constant and thereby minimize the integral over the square of the derivative of the curvature. Because of these appealing properties, the MVC fairness measure was chosen.

@@

				The first problem we face is that polynomials will in general not be in natural parametrization (that is, a parametrization where the parameter \(t\) is equal to the covered arc length of the curve up to the point \(\apply{p}{t}\)). Since we chose to give the position of a specification item as a fraction of the curve's total arc length, it would be very advantageous to have a parametrization where the parameter is at least proportional to the covered arc length of the curve. This property is equivalent to the speed of the curve being constant and has additional useful consequences, such as the second derivative of the curve being orthogonal to the first derivative. Thus, one of the objectives of the optimization is to make sure that the speed of the curve is constant.
